Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-03-07 Milestone_1/01-chap1.Rmd
add milestone report

# Introduction {#ch:intro}

## Is what we see really there?

## Artificial intelligence: Four Approaches {#se:ai}

**Artificial intelligence** (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence. As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:

With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life. This is largely as a consequence of the huge investment in AI industry by the financial market in recent years. Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015]. Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing. AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players [@silver_general_2018].

@@

Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework. @donoho_50_2017 in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001). In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions. Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable. @breiman_statistical_2001 suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function $\boldsymbol{f}(\boldsymbol{x})$ to describe the relationship between the features $\boldsymbol{x}$ and the responses $\boldsymbol{y}$.

@@

Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model. Two common types of model defects were created and visualized using residuals vs. fitted values plot. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.

## Structure of the paper

...

# Human Subject Exeriments

@@

## Cubic Model

@@

## Heteroskedasticity Model

@@

## Distribution of regressors

@@

## Experiment I and II

@@

## $P$-value and Power Estimation of Visual Test Allowed for Multiple Selections

@@

## Results

@@

### Power Estimation

@@

### Effect of Factors on Power of the Visual Test

@@

# Computer Vision Model

Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings. However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields. To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.

@@

## Automatic visual inference system

For an automatic visual inference system, there are several possibilities of the type of inputs.

@@

@@

## Model Architecture

@@

## Performance

@@

>>>>

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

# Background and Motivation

<!-- ## Motivation -->

<!-- Have you doubted what you found from a data plot was really there? Or have you had reservations about visual discoveries because they might be illusions? Data plots are prevalent in every field of nature and social science as it is one of the most recognized data visualization techniques. Without data plots, trends and patterns embed in the data would not be easy to discover. However, data plots are often over- or under-interpreted by human readers which leads to unsecured findings. This is unsatisfactory as confirmatory research is the foundation of science. -->

<!-- Visual inference, an inferential framework developed by @buja_statistical_2009, provides an option to test the visual findings, which helps solidify our confidence in using data plots. But one of the drawbacks of this framework is it requires human evaluation of data plots. Such requirement is unattainable when the number of data plots need for testing is large. Even if it can be fulfilled, the cost of labour would be exceedingly high. -->

<!-- ## Research Problem -->

<!-- The main objective of this research is to build an automatic visual inference system that can evaluate data plots. We aim to apply this automatic system on predictive model diagnostics because one of the scenarios best fitted for visual inference is where there does not exist a conventional test, and flexible predictive models which adopt machine learning algorithms do not generally have a conventional hypothesis testing procedure.  -->

@@

@donoho_50_2017 in its summary of data science stated that the concept of the predictive modelling culture could be traced back to an article written by Breiman (2001). In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.

@@

# Structure of the Thesis (need disscussion)

An overview of the thesis, including some brief detail of the proposed content of all chapters. This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions

1. Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
2. ...
3. ...

# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics

This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.

## Human Subject Exeriments

@@

### Cubic Model

@@

### Heteroskedasticity Model

@@

### Distribution of regressors

@@

### Experiment I and II

@@

### $P$-value and Power Estimation of Visual Test Allowed for Multiple Selections

@@

### Results

@@

#### Power Estimation

@@

#### Effect of Factors on Power of the Visual Test

@@

## Automatic Visual Inference System

Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings. However, the lack of scalability of this method limits its application in wider fields. The need of automatic visual inference system is auto data analysis and machine learning system.

AI agents which can read data plot and evaluate lineups could

@@

### Automatic visual inference system

For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.

single image file
multiple image files

fitted model

@@

### Artificial Intelligence: Four Approaches {#se:ai}

Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated. This brings us to one of the popular filed of study - **Artificial Intelligence** (AI). AI is the field of research concerned with understanding and building machines who can demonstrate intelligence. As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:

@@

### Model Architecture

@@

### Performance

@@

# Timeline (need discussion)

### 2022-03-07 Milestone_1/_book/thesis.tex
add milestone report

\author{Weihao Li}
\title{Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics}
\degrees{B.Comm. (Hons), Monash University}
\def\degreetitle{Doctor of Philosophy}
% Add subject and keywords below
\hypersetup{
     %pdfsubject={The Subject},
     %pdfkeywords={Some Keywords},
     pdfauthor={Weihao Li},
     pdftitle={Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics},
     pdfproducer={Bookdown with LaTeX}
}

@@

\hypertarget{copyright-notice}{%
\chapter*{Copyright notice}\label{copyright-notice}}
\addcontentsline{toc}{chapter}{Copyright notice}

\emph{(Choose one of the following notices.)}

\emph{(Notice 1)}

\textcopyright { } \authorname~(\number\the\year).

\emph{The second notice certifies the appropriate use of any third-party material in the thesis. Students choosing to deposit their thesis into the restricted access section of the repository are not required to complete Notice 2.}

\emph{(Notice 2)}

\textcopyright { } \authorname~(\number\the\year).

I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.

\newpage

\hypertarget{abstract}{%
\chapter*{Abstract}\label{abstract}}
\addcontentsline{toc}{chapter}{Abstract}

The abstract should outline the main approach and findings of the thesis and must not be more than 500 words.

\newpage

\hypertarget{declaration}{%
\chapter*{Declaration}\label{declaration}}
\addcontentsline{toc}{chapter}{Declaration}

\emph{(Standard thesis)}

This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

\emph{(Thesis including published works declaration)}

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes (insert number) original papers published in peer reviewed journals and (insert number) submitted publications. The core theme of the thesis is (insert theme). The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the (insert name of academic unit) under the supervision of (insert name of supervisor).

(The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.) Remove this paragraph for theses with sole-authored work

In the case of (insert chapter numbers) my contribution to the work involved the following:

I have / have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

\begin{table}
\centering\footnotesize\tabcolsep=0.12cm
\begin{tabular}{|p{1.2cm}|p{2cm}|p{1.8cm}|p{3.4cm}|p{3.5cm}|p{1.5cm}|}
\hline
\RaggedRight\textbf{Thesis chapter}  & 
\RaggedRight\textbf{Publication title}  & 
\RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  &  
\RaggedRight\textbf{Nature and \% of student contribution} & 
\RaggedRight\textbf{Co-author name(s), nature and \% of co-author’s contribution} &  \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ 
\hline
2 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
\hline
3 & xx & xx  & xx & xx  & \multicolumn{1}{c|}{N}   \\
\hline
4 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}    \\
\hline
5 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
\hline
\end{tabular}
\end{table}

\textbf{Student name:} \authorname

\textbf{Student signature:}

\textbf{Date:}

\hypertarget{acknowledgements}{%
\chapter*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my pet goldfish for \dots

\hypertarget{preface}{%
\chapter*{Preface}\label{preface}}
\addcontentsline{toc}{chapter}{Preface}

The material in Chapter \ref{ch:intro} has been submitted to the journal \emph{Journal of Impossible Results} for possible publication.

The contribution in Chapter \ref{ch:litreview} of this thesis was presented in the International Symposium on Nonsense held in Dublin, Ireland, in July 2015.

@@

\hypertarget{ch:intro}{%
\chapter{Introduction}\label{ch:intro}}

\hypertarget{is-what-we-see-really-there}{%
\section{Is what we see really there?}\label{is-what-we-see-really-there}}

\hypertarget{se:ai}{%
\section{Artificial intelligence: Four Approaches}\label{se:ai}}

\textbf{Artificial intelligence} (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence. As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:

With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life. This is largely as a consequence of the huge investment in AI industry by the financial market in recent years. Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history \autocite{jordan_machine_2015}. Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing. AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players \autocite{silver_general_2018}.

@@

Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework. \textcite{donoho_50_2017} in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001). In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions. Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable. \textcite{breiman_statistical_2001} suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function \(\boldsymbol{f}(\boldsymbol{x})\) to describe the relationship between the features \(\boldsymbol{x}\) and the responses \(\boldsymbol{y}\).

@@

Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model. Two common types of model defects were created and visualized using residuals vs.~fitted values plot. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.

\hypertarget{structure-of-the-paper}{%
\section{Structure of the paper}\label{structure-of-the-paper}}

\ldots{}

\hypertarget{human-subject-exeriments}{%
\chapter{Human Subject Exeriments}\label{human-subject-exeriments}}

@@

\hypertarget{cubic-model}{%
\section{Cubic Model}\label{cubic-model}}

@@

\hypertarget{heteroskedasticity-model}{%
\section{Heteroskedasticity Model}\label{heteroskedasticity-model}}

@@

\hypertarget{distribution-of-regressors}{%
\section{Distribution of regressors}\label{distribution-of-regressors}}

@@

\hypertarget{experiment-i-and-ii}{%
\section{Experiment I and II}\label{experiment-i-and-ii}}

@@

\hypertarget{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}{%
\section{\texorpdfstring{\(P\)-value and Power Estimation of Visual Test Allowed for Multiple Selections}{P-value and Power Estimation of Visual Test Allowed for Multiple Selections}}\label{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

@@

\hypertarget{power-estimation}{%
\subsection{Power Estimation}\label{power-estimation}}

@@

\hypertarget{effect-of-factors-on-power-of-the-visual-test}{%
\subsection{Effect of Factors on Power of the Visual Test}\label{effect-of-factors-on-power-of-the-visual-test}}

@@

\hypertarget{computer-vision-model}{%
\chapter{Computer Vision Model}\label{computer-vision-model}}

Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings. However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields. To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.

@@

\hypertarget{automatic-visual-inference-system}{%
\section{Automatic visual inference system}\label{automatic-visual-inference-system}}

For an automatic visual inference system, there are several possibilities of the type of inputs.

@@

@@

\hypertarget{model-architecture}{%
\section{Model Architecture}\label{model-architecture}}

@@

\hypertarget{performance}{%
\section{Performance}\label{performance}}

@@

\appendix

\hypertarget{additional-stuff}{%
\chapter{Additional stuff}\label{additional-stuff}}

You might put some computer output here, or maybe additional tables.

Note that line 5 must appear before your first appendix. But other appendices can just start like any other chapter.

>>>>

\author{Weihao Li}
\title{Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics}
\degrees{B.Comm. (Hons), Monash University}
\def\degreetitle{PhD}
% Add subject and keywords below
\hypersetup{
     %pdfsubject={The Subject},
     %pdfkeywords={Some Keywords},
     pdfauthor={Weihao Li},
     pdftitle={Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics},
     pdfproducer={Bookdown with LaTeX}
}

@@

@@

\hypertarget{background-and-motivation}{%
\chapter{Background and Motivation}\label{background-and-motivation}}

@@

\textcite{donoho_50_2017} in its summary of data science stated that the concept of the predictive modelling culture could be traced back to an article written by Breiman (2001). In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.

@@

\hypertarget{structure-of-the-thesis-need-disscussion}{%
\chapter{Structure of the Thesis (need disscussion)}\label{structure-of-the-thesis-need-disscussion}}

An overview of the thesis, including some brief detail of the proposed content of all chapters. This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
\item
  \ldots{}
\item
  \ldots{}
\end{enumerate}

\hypertarget{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}{%
\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}}

This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.

\hypertarget{human-subject-exeriments}{%
\section{Human Subject Exeriments}\label{human-subject-exeriments}}

@@

\hypertarget{cubic-model}{%
\subsection{Cubic Model}\label{cubic-model}}

@@

\hypertarget{heteroskedasticity-model}{%
\subsection{Heteroskedasticity Model}\label{heteroskedasticity-model}}

@@

\hypertarget{distribution-of-regressors}{%
\subsection{Distribution of regressors}\label{distribution-of-regressors}}

@@

\hypertarget{experiment-i-and-ii}{%
\subsection{Experiment I and II}\label{experiment-i-and-ii}}

@@

\hypertarget{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}{%
\subsection{\texorpdfstring{\(P\)-value and Power Estimation of Visual Test Allowed for Multiple Selections}{P-value and Power Estimation of Visual Test Allowed for Multiple Selections}}\label{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

@@

\hypertarget{power-estimation}{%
\subsubsection{Power Estimation}\label{power-estimation}}

@@

\hypertarget{effect-of-factors-on-power-of-the-visual-test}{%
\subsubsection{Effect of Factors on Power of the Visual Test}\label{effect-of-factors-on-power-of-the-visual-test}}

@@

\hypertarget{automatic-visual-inference-system}{%
\section{Automatic Visual Inference System}\label{automatic-visual-inference-system}}

Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings. However, the lack of scalability of this method limits its application in wider fields. The need of automatic visual inference system is auto data analysis and machine learning system.

AI agents which can read data plot and evaluate lineups could

@@

\hypertarget{automatic-visual-inference-system-1}{%
\subsection{Automatic visual inference system}\label{automatic-visual-inference-system-1}}

For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.

single image file
multiple image files

fitted model

@@

\hypertarget{se:ai}{%
\subsection{Artificial Intelligence: Four Approaches}\label{se:ai}}

Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated. This brings us to one of the popular filed of study - \textbf{Artificial Intelligence} (AI). AI is the field of research concerned with understanding and building machines who can demonstrate intelligence. As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:

@@

\hypertarget{model-architecture}{%
\subsection{Model Architecture}\label{model-architecture}}

@@

\hypertarget{performance}{%
\subsection{Performance}\label{performance}}

@@

\hypertarget{timeline-need-discussion}{%
\chapter{Timeline (need discussion)}\label{timeline-need-discussion}}

### 2022-03-07 Milestone_1/index.Rmd
add milestone report

---
title: 'Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics'
degreetype: 'Doctor of Philosophy'
author: 'Weihao Li'
degrees: 'B.Comm. (Hons), Monash University'
output: bookdown::pdf_book
site: bookdown::bookdown_site
link-citations: yes
knit: "bookdown::render_book"
---

>>>>

---
title: 'Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics'
degreetype: 'PhD'
author: 'Weihao Li'
degrees: 'B.Comm. (Hons), Monash University'
output: bookdown::pdf_book
site: bookdown::bookdown_site
link-citations: yes
knit: "bookdown::render_book"
---
